---
title: "Smoking in several social dimensions"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---



## Setup

### Load packages

```{r load-packages, message = FALSE}
pacman::p_load(dplyr, ggplot2, kableExtra)
```

### Load data


```{r load-data}
setwd("H:\\LITERATURA_i_POBRANE\\R_kurs\\Cursera R\\week1_statysitical")
load("brfss2013.gz")
```



* * *

## Part 1: Data
    
#  relationship between age and smoking  -----------------

    



* * *

## Part 2: Research questions

# ASTMA i HAVY smookers

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r fig.width=13, fig.height=4, echo=FALSE}
df1<- brfss2013 %>% 
    select(X_ageg5yr, smokday2) %>% 
    filter(!is.na(smokday2)) %>% 
    mutate(smoker = ifelse(smokday2 == 'Not at all', 'No','Yes')) 


ggplot()  + 
    geom_bar(data = df1 ,aes(x = X_ageg5yr,  fill =  smoker), position = "dodge" ) +
    labs(x = "",
         y = "Number of smokers",
         title = " Relationship between age and smoking") +
    theme_minimal()+
    theme(legend.title=element_blank(),
          plot.title = element_text(hjust = 0.5))

```



**Research quesion 2:**

```{r}
df2 <- brfss2013 %>% 
    select(smoke100, asattack) %>% 
    na.omit() %>% 
    group_by(asattack, smoke100 ) %>% 
    rename(`Asthma attack during past 12 month` = asattack, `Smoked At Least 100 Cigarettes` = smoke100) %>% 
    summarise(`The numbers` = n()) %>% 
    kable() %>%
    kable_styling()
df2



```



**Research quesion 3:**
```{r}
df3 <- brfss2013 %>% 
    select(smokday2, exerhmm1, income2 ) %>% 
    na.omit() %>% 
    mutate( activity_level = cut(exerhmm1, breaks=c(1, 30, 60, 120, max(exerhmm1)), include.lowest = TRUE)) %>% 
    mutate(smoker = ifelse(smokday2 == 'Not at all', 'No','Yes')) %>% 
    select(income2, activity_level, smoker ) %>% 
    group_by(activity_level, income2, smoker) %>% 
    summarise( L = n()) %>% 
    group_by(activity_level, income2) %>% 
    mutate( L1 = sum(L)) %>% 
    filter(smoker == 'Yes') %>% 
    summarise(smokers_percentage = round( L / L1 * 100, 0 )) %>% 
    arrange(income2, activity_level)



ggplot(data = df3) +
    geom_point(aes(x = activity_level, y = income2,  size = smokers_percentage)) +
    theme_minimal()+
    labs(x = "Activity time per week (in minutes)",
         y = "Incame",
         title = " Relationship between \n smoking, income and activity") +
    # theme_minimal()+
    theme( plot.title = element_text(hjust = 0.5)) +
    scale_size(name = "Smokers percentage",
               range = c(0, 10))
 
```

